---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tswge)
library(dplyr)
library(vars)
```

```{r}
la_mort <- read.csv("https://raw.githubusercontent.com/boneeyah/DS6373/main/la_cmort_study%20(2).csv")
```

first we plot the realizations
The patterns here show evidence of similar cyclyc behaviors which appear to be phase shifted for temp, and in phase for part and cmort. They appear to have similar structures, with no major concerns, although the peak at around 150 for cmort is higher than the rest.
Also noting, the period appears to be around 50 for each, which would make it about a yearly period since these are weekly time series.

```{r}
plot.ts(la_mort[,2:4])
```
Although a linear trend does not look like a good model for the data, we will fit a GLP as a baseline


```{r}
lfit <- lm(cmort~temp+part,data = la_mort)
la_temp <- la_mort$temp
la_part <- la_mort$part
ar_p <- aic.wge(lfit$residuals,p=0:10,q=0) # selects for AR(4)
res_fit <- arima(la_mort$cmort,order = c(ar_p$p,0,0),xreg = cbind(la_temp,la_part))
res_fit
```
As expected based on the phase shift, the temperature is not significant (s.e is actually larger than the coefficient), while la_part appears to be significant.

There are 2 autocorrelations above the confidence limits, and the ljung-box test returns .08 and .04 for K=24 and K=48 respectively, meaning the data is not properly modeled, will fit a better model with the 
```{r}
plotts.sample.wge(res_fit$residuals,arlimits = TRUE)
ljung.wge(res_fit$residuals,p=4,K=24)
ljung.wge(res_fit$residuals,p=4,K=48)
```
```{r}
ccf(la_mort$part,la_mort$cmort,main='Mortality-Pollution Correlation')
ccf(la_mort$temp,la_mort$cmort,main='Mortality-Temperature Correlation')
```
```{r}
la_temp1 <- lag(la_mort$temp,1)
la_part7 <- lag(la_mort$part,7)
```

```{r}
lfit1 <- lm(la_mort$cmort~la_temp1 + la_part7)
ar_p1 <- aic.wge(lfit1$residuals,p=0:10,q=0)
res_fit1 <- arima(la_mort$cmort,order = c(ar_p1$p,0,0),xreg=cbind(la_temp1,la_part7))
res_fit1
```
for the new model, intercept, temperature with lag 1 and pollution with lag 7 all appear to be significant.
The residuals plot look properly whitened, with no lags above the confidence limits. Additonaly the ljung-box test for the residuals with K=24 and K=48 yieleded p-values of .83 and .79 respectively providing additional evidence, suggesting the model has approprietely handled the data
```{r}
plotts.sample.wge(res_fit1$residuals[8:508],arlimits = TRUE)
ljung.wge(res_fit1$residuals,p=4,K=24)
ljung.wge(res_fit1$residuals,p=4,K=48)
```
Proceeding with the VAR modeling 
```{r}
X <- la_mort[1:456,2:4]
VARselect(X,lag.max = 20,type = "const") #chooses lag of 9 for AIC and 2 for BIC
```
```{r}
varfit <- VAR(X,p = 2,type = "both")
varfit1 <- VAR(X,p=9,type = "both")
summary(varfit)
summary(varfit1)
```

```{r}
lpreds <- predict(res_fit1,newxreg = data.frame(la_temp1 = la_temp1[457:508], la_part7 = la_part7[457:508]))
varpreds <- predict(varfit,n.ahead = 52)
varpreds1 <- predict(varfit1,n.ahead = 52)
```

```{r}
plot(la_mort$Week,la_mort$cmort,type='l',main='MLR model with AR(9)')
points(la_mort$Week[457:508],lpreds$pred,type = 'l',col='red',lwd=2)
plot(la_mort$Week,la_mort$cmort,type='l',main='VAR(2) model')
points(la_mort$Week[457:508],varpreds$fcst$cmort[,1],type = 'l',col='red',lwd=2)
plot(la_mort$Week,la_mort$cmort,type='l',main='VAR(9) model')
points(la_mort$Week[457:508],varpreds1$fcst$cmort[,1],type = 'l',col='red',lwd=2)
```
```{r}
l_ase <- mean((la_mort$cmort[457:508] - lpreds$pred)^2)
var_ase <- mean((la_mort$cmort[457:508] - varpreds$fcst$cmort[,1])^2)
var1_ase <- mean((la_mort$cmort[457:508] - varpreds1$fcst$cmort[,1])^2)

cat('MLR ASE = ',l_ase)
cat('\nVAR AR(2) ASE = ',var_ase)
cat('\nVAR AR(9) ASE =', var1_ase)
```
```{r}
business <- read.csv("https://raw.githubusercontent.com/boneeyah/DS6373/main/businesssales.csv")
```

```{r}
plot.ts(business[,2:5])
```
```{r}
ccf(business$ad_tv,business$sales)
ccf(business$ad_online,business$sales)
ccf(business$discount,business$sales)
```

```{r}
ad_tv1 <- lag(business$ad_tv,1)
ad_online1 <- lag(business$ad_online,1)
#discount <- busi$discount
discount4 <- lag(business$discount,4)
busi <- data.frame(sales = busi$sales,ad_tv1,ad_online1,discount4 = discount4)
busi1 <- busi[0:95,]
mlrfit <- lm(sales~ad_tv1+ad_online1+discount4,data=busi1)
aic.wge(mlrfit$residuals,p=0:10,q=0)
fit = arima(busi1$sales,order = c(7,0,0),xreg = cbind(busi1$ad_tv1,busi1$ad_online1,busi1$discount4))
fit
```
```{r}
plotts.sample.wge(fit$residuals[5:95],arlimits = TRUE)
ljung.wge(fit$residuals[5:95])
ljung.wge(fit$residuals[5:95],K=48)
```
```{r}
VARselect(business[,2:5],lag.max = 10,type = "const")
VAR2_fit <- VAR(business[0:95,2:5],p=2,type = "both")
VAR3_fit <- VAR(business[0:95,2:5],p=3,type = "both")
```

```{r}
mlr_pred <- predict(fit,newxreg = data.frame(busi$ad_tv1[96:100],busi$ad_online1[96:100],busi$discount4[96:100]))
var2_pred <- predict(VAR2_fit,n.ahead = 25)
var3_pred  <- predict(VAR3_fit,n.ahead = 25)
```
```{r}
x <- 96:100
plot(1:100,busi$sales,type='l')
points(x,mlr_pred$pred,col='red',type = 'o')
```
```{r}
plot(1:100,busi$sales,type='l',xlim=c(1,120))
points(96:120,var2_pred$fcst$sales[,1],col='blue',type = 'o')
```

```{r}
plot(1:100,busi$sales,type='l',xlim=c(0,120))
points(96:120,var3_pred$fcst$sales[,1],col='pink',type='o')
```

```{r}
mlr_ase <- mean((busi$sales[96:100] - mlr_pred$pred)^2)
var2_ase <- mean((busi$sales[96:100] - var2_pred$fcst$sales[,1])^2)
var3_ase <- mean((busi$sales[96:100] - var3_pred$fcst$sales[,1])^2)
```

