---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tswge)
```

```{r}
accu <- read.csv("https://raw.githubusercontent.com/boneeyah/DS6373/main/accuspike%20(2).csv")
x <- accu$Active.Users[0:183]
```

```{r}
plotts.wge(x)
plotts.sample.wge(x)
```
The plots above do not show evidence of seasonality. The ACF plot does not show much evidence of non-stationary behavior.
But will look at the factor table to see how strong root with frequency of 0 is
```{r}
est.arma.wge(x,p=13)
```
As expected from the specntral density plot, there appears to be a root at 0, but it is not dominant. It's associated absoulte reciprocal is .93, which means it is not particularly close to the unit circle, meaning the frequency of 0 is not associated with non-stationary behavior.
We will proceed with a stationary model

```{r}
x <- artrans.wge(x,phi.tr = c(rep(0,6),1))
aic5.wge(x,p=0:14,q=0:2,type = 'aic')
```
Will go with the 14,0 model since this is daily data, biweekly patterns seem more plausible than 10 or 12 day patterns (though they are also possible).

```{r}
est14 <- est.arma.wge(x,p=14,q=1)
```

```{r}
fore.arima.wge(accu$Active.Users[0:183],phi=est14$phi,theta = est14$theta,n.ahead = 14,lastn = TRUE,s=7)
```

