---
title: "FLS 11"
output: html_notebook
---
1. Plot the latest yearly sunspot data
```{r}
sun_spots <- read.csv("https://raw.githubusercontent.com/boneeyah/DS6373/main/SN_y_tot_V2.0.csv",sep = ";",header = FALSE)
x <- sun_spots$V2
plotts.wge(x)
```
2. Comment on its stationarity:
We see some evidence against stationarity. More specifically, the assumption of the mean sunsposts being independent of time appears like it might be violated.

Based on the realization, the acf plot and the spectral density plot, there is not strong evidence of wandering. There is a peak at frequency of 0, but it is not the dominant behavior, so it does not appear to correspond to a non-stationary root on the unit circle.
```{r}
plotts.sample.wge(x)
```
3. Use aic5.wge to estimate the p and the q of the model.

```{r}
aic5.wge(x,p=0:14, q=0:8,type = 'aic')

aic5.wge(x,p=0:14,q=0:8,type = 'bic')
```
4. Fit the model using your model identificiation (p and q)
```{r}
arma3.4 <- est.arma.wge(x,p=3,q=4)
arma12.3 <- est.arma.wge(x,p=12,q=3)

fore3.4 <- fore.arma.wge(x,phi = arma3.4$phi,theta = arma3.4$theta,n.ahead = 15,lastn = TRUE,limits = FALSE)
fore12.3 <- fore.arma.wge(x,phi = arma12.3$phi,theta = arma12.3$theta,n.ahead = 15,lastn = TRUE,limits = FALSE)
```
5. Use this model to generate an ASE from forecasting the last 15 years of sunspot data.
```{r}
ASE <- mean((x[309:323] - fore12.3$f)^2)

roll.win.rmse.wge(x,horizon = 15,phi = arma12.3$phi,theta = arma12.3$theta)
```
6. Now fit a seasonal model to the data.
```{r}
est.arma.wge(x,p=16)
factor.wge(phi = c(rep(0,10),1))

#we can see that it matches a seasonality of 11 fairly well
d11 <- artrans.wge(x,phi.tr = c(rep(0,10),1))
#acfs and realization plot do not look like white noise, so we will still try to find and fit an ar model
parzen.wge(d11)

aic5.wge(d11,p=0:12,q=0:8,type = 'aic')
d11.11.3 <- est.arma.wge(d11,p=11,q=3)
plotts.wge(d11.11.3$res)
acf(d11.11.3$res)

fore11.11.3 <- fore.arima.wge(x,phi = d11.11.3$phi,theta = d11.11.3$theta,s=11,n.ahead = 15,lastn = TRUE,limits = FALSE)

ASE2 <- mean((fore11.11.3$f - x[309:323])^2)
roll.win.rmse.wge(x,horizon = 15,s=11,phi = d11.11.3$phi,theta = d11.11.3$theta)
```

